---
title: "Examples in other dimensions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples in other dimensions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{ggplot2,patchwork}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  fig.asp = 0.5,
  dpi = 200
)
```

This vignette contains a number of examples on how to use `clugenr` in 1D and
5D. Examples require the following setup code:

```{r setup, message=FALSE, warning=FALSE}
# Load the clugenr library
library(clugenr)

# Load functions for plotting examples
source("plot_examples_1d.R", local = knitr::knit_global())
source("plot_examples_nd.R", local = knitr::knit_global())

# Keep examples reproducible in newer R versions
RNGversion("3.6.0")
```

## Basic 1D example with density plot

The 1D examples were plotted with the `plot_examples_1d()` function available
[here](https://github.com/clugen/clugenr/blob/main/vignettes/plot_examples_1d.R).

```{r}
seed <- 222
```

```{r}
# Custom proj_dist_fn: point projections placed using the Weibull distribution
proj_wbull <- function(len, n) rweibull(n, shape = 1.5, scale = len / 2) - len / 2
```

```{r}
e079 <- clugen(1, 3, 2000, 1, 0, 10, 6, 1.5, 0, seed = seed)
e080 <- clugen(1, 3, 2000, 1, 0, 10, 6, 1.5, 0, seed = seed, proj_dist_fn = "unif")
e081 <- clugen(1, 3, 2000, 1, 0, 10, 6, 1.5, 0, seed = seed, proj_dist_fn = proj_wbull)
```

```{r}
plot_examples_1d(list(e = e079, t = "e079: proj_dist_fn='norm'"),
                 list(e = e080, t = "e080: proj_dist_fn='unif'"),
                 list(e = e081, t = "e081: custom proj_dist_fn (Weibull)"),
                 pmargin = 0.07,
                 ymax = 0.575)
```

## 5D example with default optional arguments

The 5D examples were plotted with the `plot_examples_nd()` function available
[here](https://github.com/clugen/clugenr/blob/main/vignettes/plot_examples_nd.R).

```{r}
nd <- 5
seed <- 321
```

```{r}
e082 <- clugen(nd, 6, 1500, c(1, 1, 0.5, 0, 0), pi / 16, rep.int(30, nd), 30, 4, 3,
              seed = seed)
```

```{r, fig.asp = 1}
plot_examples_nd(e082, "e082: 5D with optional parameters set to defaults")
```

## 5D example with `proj_dist_fn = "unif"` and `point_dist_fn = "n"`

```{r}
nd <- 5
seed <- 123
```

```{r}
e083 <- clugen(nd, 6, 1500, c(0.1, 0.3, 0.5, 0.3, 0.1), pi / 12, rep.int(30, nd),
              35, 5, 3.5, seed = seed,
              proj_dist_fn = "unif", point_dist_fn = "n")
```

```{r, fig.asp = 1}
plot_examples_nd(e083, "e083: 5D with proj_dist_fn='unif' and point_dist_fn='n'")
```

