---
title: "Examples in 2D"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples in 2D}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  fig.asp = 0.5,
  dpi = 200
)
```

This vignette contains a number of examples on how to use `clugenr` in 2D.
Examples require the following setup code:

```{r setup, message=FALSE, warning=FALSE}
library(clugenr)   # The clugenr library
library(ggplot2)   # For plotting
library(patchwork) # For combining plots

# Keep examples reproducible in newer R versions 
RNGversion("3.6.0") 

# Function for plotting a series of related examples
plot_examples <- function(...) {
  pmargin <- 0.1
  ets <- list(...)
  xmax <- max(sapply(ets, function(et) max(et$e$points[, 1])))
  xmin <- min(sapply(ets, function(et) min(et$e$points[, 1])))
  ymax <- max(sapply(ets, function(et) max(et$e$points[, 2])))
  ymin <- min(sapply(ets, function(et) min(et$e$points[, 2])))
  xd <- pmargin * abs(xmax - xmin)
  yd <- pmargin * abs(ymax - ymin)
  xmax <- xmax + xd
  xmin <- xmin - xd
  ymax <- ymax + yd
  ymin <- ymin - yd
  plts <- lapply(
    ets,
    function(et) {
      e <- et$e
      t <- et$t
      ggplot(NULL, aes(x = e$points[, 1], y = e$points[, 2])) +
      geom_point(shape = 21, colour = "black", stroke = 0.1,
                 aes(fill = factor(e$point_clusters))) +
      xlab(NULL) + ylab(NULL) + ggtitle(t) +
      theme(legend.position = "none",
            plot.title = element_text(size = rel(0.9))) +
      coord_fixed(xlim = c(xmin, xmax), ylim = c(ymin, ymax))
    })
  wrap_plots(plts)
}
```

## Manipulating the direction of cluster-supporting lines

### Using the `direction` parameter

```{r}
seed <- 123
```

```{r}
set.seed(seed)
e01 <- clugen(2, 4, 200, c(1, 0), 0, c(10, 10), 10, 1.5, 0.5)
set.seed(seed)
e02 <- clugen(2, 4, 200, c(1, 1), 0, c(10, 10), 10, 1.5, 0.5)
set.seed(seed)
e03 <- clugen(2, 4, 200, c(0, 1), 0, c(10, 10), 10, 1.5, 0.5)

plot_examples(list(e = e01, t = "e01: direction = [1, 0]"),
              list(e = e02, t = "e02: direction = [1, 1]"),
              list(e = e03, t = "e03: direction = [0, 1]"))
```

### Changing the `angle_disp` parameter and using a custom `angle_deltas_fn` function

```{r}
seed <- 123

# Custom angle_deltas function: arbitrarily rotate some clusters by 90 degrees
angdel_90_fn <- function(nclu, astd) sample(c(0, pi / 2), nclu, replace = T)
```

```{r}
set.seed(seed)
e04 <- clugen(2, 6, 500, c(1, 0), 0, c(10, 10), 10, 1.5, 0.5)
set.seed(seed)
e05 <- clugen(2, 6, 500, c(1, 0), pi / 8, c(10, 10), 10, 1.5, 0.5)
set.seed(seed)
e06 <- clugen(2, 6, 500, c(1, 0), 0, c(10, 10), 10, 1.5, 0.5,
              angle_deltas_fn = angdel_90_fn)

plot_examples(list(e = e04, t = "e04: angle_disp = 0"),
              list(e = e05, t = "e05: angle_disp = π/8"),
              list(e = e06, t = "e06: custom angle_deltas function"))
```

## Manipulating the length of cluster-supporting lines

### Using the `llength` parameter

```{r}
seed <- 321
```

```{r}
set.seed(seed)
e07 <- clugen(2, 5, 800, c(1, 0), pi / 10, c(10, 10), 0, 0, 0.5,
              point_dist_fn = "n")
set.seed(seed)
e08 <- clugen(2, 5, 800, c(1, 0), pi / 10, c(10, 10), 10, 0, 0.5,
              point_dist_fn = "n")
set.seed(seed)
e09 <- clugen(2, 5, 800, c(1, 0), pi / 10, c(10, 10), 30, 0, 0.5,
              point_dist_fn = "n")

plot_examples(list(e = e07, t = "e07: llength = 0"),
              list(e = e08, t = "e08: llength = 10"),
              list(e = e09, t = "e09: llength = 30"))
```

### Changing the `llength_disp` parameter and using a custom `llengths_fn` function

```{r}
seed <- 765

# Custom llengths function: line lengths grow for each new cluster
llen_grow_fn <- function(nclu, llen, llenstd) llen * 0:(nclu - 1) + rnorm(nclu, sd = llenstd)
```

```{r}
set.seed(seed)
e10 <- clugen(2, 5, 800, c(1, 0), pi / 10, c(10, 10), 15,  0.0, 0.5,
              point_dist_fn = "n")
set.seed(seed)
e11 <- clugen(2, 5, 800, c(1, 0), pi / 10, c(10, 10), 15, 10.0, 0.5,
              point_dist_fn = "n")
set.seed(seed)
e12 <- clugen(2, 5, 800, c(1, 0), pi / 10, c(10, 10), 10,  0.1, 0.5,
              llengths_fn = llen_grow_fn, point_dist_fn = "n")

plot_examples(list(e = e10, t = "e10: llength_disp = 0.0"),
              list(e = e11, t = "e11: llength_disp = 5.0"),
              list(e = e12, t = "e12: custom llengths function"))
```

## Manipulating relative cluster positions

### Using the `cluster_sep` parameter

```{r}
seed <- 222
```

```{r}
set.seed(seed)
e13 <- clugen(2, 8, 1000, c(1, 1), pi / 4, c(10, 10), 15, 2, 4)
set.seed(seed)
e14 <- clugen(2, 8, 1000, c(1, 1), pi / 4, c(30, 10), 15, 2, 4)
set.seed(seed)
e15 <- clugen(2, 8, 1000, c(1, 1), pi / 4, c(10, 30), 15, 2, 4)

plot_examples(list(e = e13, t = "e13: cluster_sep = [10, 10]"),
              list(e = e14, t = "e14: cluster_sep = [30, 10]"),
              list(e = e15, t = "e15: cluster_sep = [10, 30]"))
```

### Changing the `cluster_offset` parameter and using a custom `clucenters_fn` function

```{r}
seed <- 222

# Custom clucenters function: places clusters in a diagonal
centers_diag <- function(nclu, csep, coff) {
  matrix(1, nrow = nclu, ncol = length(csep)) * (1:nclu * max(csep)) +
    rep(coff, each = nclu)
}
```

```{r}
set.seed(seed)
e16 <- clugen(2, 8, 1000, c(1, 1), pi / 4, c(10, 10), 10, 2, 2.5)
set.seed(seed)
e17 <- clugen(2, 8, 1000, c(1, 1), pi / 4, c(10, 10), 10, 2, 2.5,
              cluster_offset = c(20, -20))
set.seed(seed)
e18 <- clugen(2, 8, 1000, c(1, 1), pi / 4, c(10, 10), 10, 2, 2.5, 
              cluster_offset = c(-50, -50), clucenters_fn = centers_diag)

plot_examples(list(e = e16, t = "e16: default"),
              list(e = e17, t = "e17: cluster_offset = [20, -20]"),
              list(e = e18, t = "e18: custom clucenters function"))
```

## Lateral dispersion and placement of point projections on the line

```{r}
seed <- 234
```

### Normal projection placement (default): `proj_dist_fn = "norm"`

```{r}
set.seed(seed)
e19 <- clugen(2, 4, 1000, c(1, 0), pi / 2, c(20, 20), 13, 2, 0.0)
set.seed(seed)
e20 <- clugen(2, 4, 1000, c(1, 0), pi / 2, c(20, 20), 13, 2, 1.0)
set.seed(seed)
e21 <- clugen(2, 4, 1000, c(1, 0), pi / 2, c(20, 20), 13, 2, 3.0)

plot_examples(list(e = e19, t = "e19: lateral_disp = 0"),
              list(e = e20, t = "e20: lateral_disp = 1"),
              list(e = e21, t = "e21: lateral_disp = 3"))
```

### Uniform projection placement: `proj_dist_fn = "unif"`

```{r}
set.seed(seed)
e22 <- clugen(2, 4, 1000, c(1, 0), pi / 2, c(20, 20), 13, 2, 0.0,
              proj_dist_fn = "unif")
set.seed(seed)
e23 <- clugen(2, 4, 1000, c(1, 0), pi / 2, c(20, 20), 13, 2, 1.0,
              proj_dist_fn = "unif")
set.seed(seed)
e24 <- clugen(2, 4, 1000, c(1, 0), pi / 2, c(20, 20), 13, 2, 3.0,
              proj_dist_fn = "unif")

plot_examples(list(e = e22, t = "e22: lateral_disp = 0"),
              list(e = e23, t = "e23: lateral_disp = 1"),
              list(e = e24, t = "e24: lateral_disp = 3"))
```

### Custom projection placement using the Beta distribution

```{r}
# Custom proj_dist_fn: point projections placed using the Beta distribution
proj_beta <- function(len, n) len * rbeta(n, 0.1, 0.1) - len / 2
```

```{r}
set.seed(seed)
e25 <- clugen(2, 4, 1000, c(1, 0), pi / 2, c(20, 20), 13, 2, 0.0,
              proj_dist_fn = proj_beta)
set.seed(seed)
e26 <- clugen(2, 4, 1000, c(1, 0), pi / 2, c(20, 20), 13, 2, 1.0,
              proj_dist_fn = proj_beta)
set.seed(seed)
e27 <- clugen(2, 4, 1000, c(1, 0), pi / 2, c(20, 20), 13, 2, 3.0,
              proj_dist_fn = proj_beta)

plot_examples(list(e = e25, t = "e25: lateral_disp = 0"),
              list(e = e26, t = "e26: lateral_disp = 1"),
              list(e = e27, t = "e27: lateral_disp = 3"))
```

## Controlling final point positions from their projections on the cluster-supporting line

```{r}
seed <- 12321

# Custom proj_dist_fn: point projections placed using the Beta distribution
proj_beta <- function(len, n) len * rbeta(n, 0.1, 0.1) - len / 2
```

### Points on hyperplane orthogonal to cluster-supporting line (default): `point_dist_fn = "n-1"`

```{r}
set.seed(seed)
e28 <- clugen(2, 5, 1500, c(1, 0), pi / 4, c(20, 20), 14, 2, 2.0)
set.seed(seed)
e29 <- clugen(2, 5, 1500, c(1, 0), pi / 4, c(20, 20), 14, 2, 2.0,
              proj_dist_fn = "unif")
set.seed(seed)
e30 <- clugen(2, 5, 1500, c(1, 0), pi / 4, c(20, 20), 14, 2, 2.0,
              proj_dist_fn = proj_beta)

plot_examples(list(e = e28, t = "e28: proj_dist_fn = 'norm' (default)"),
              list(e = e29, t = "e29: proj_dist_fn = 'unif'"),
              list(e = e30, t = "e30: custom proj_dist_fn (Beta dist.)"))
```

### Points around projection on cluster-supporting line: `point_dist_fn = "n"`

```{r}
set.seed(seed)
e31 <- clugen(2, 5, 1500, c(1, 0), pi / 4, c(20, 20), 14, 2, 2.0,
              point_dist_fn = "n")
set.seed(seed)
e32 <- clugen(2, 5, 1500, c(1, 0), pi / 4, c(20, 20), 14, 2, 2.0,
              point_dist_fn = "n", proj_dist_fn = "unif")
set.seed(seed)
e33 <- clugen(2, 5, 1500, c(1, 0), pi / 4, c(20, 20), 14, 2, 2.0,
              point_dist_fn = "n", proj_dist_fn = proj_beta)

plot_examples(list(e = e31, t = "e31: proj_dist_fn = 'norm' (default)"),
              list(e = e32, t = "e32: proj_dist_fn = 'unif'"),
              list(e = e33, t = "e33: custom proj_dist_fn (Beta dist.)"))
```

### Custom point placement using the exponential distribution

```{r}
# Custom point_dist_fn: final points placed using the Exponential distribution
clupoints_n_1_exp <- function(projs, lat_std, len, clu_dir, clu_ctr) {
    dist_exp <- function(npts, lstd) lstd * rexp(npts, rate = 2 / lstd)
    clupoints_n_1_template(projs, lat_std, clu_dir, dist_exp)
}
```

```{r}
set.seed(seed)
e34 <- clugen(2, 5, 1500, c(1, 0), pi / 4, c(20, 20), 14, 2, 2.0,
              point_dist_fn = clupoints_n_1_exp)
set.seed(seed)
e35 <- clugen(2, 5, 1500, c(1, 0), pi / 4, c(20, 20), 14, 2, 2.0,
              point_dist_fn = clupoints_n_1_exp, proj_dist_fn = "unif")
set.seed(seed)
e36 <- clugen(2, 5, 1500, c(1, 0), pi / 4, c(20, 20), 14, 2, 2.0,
              point_dist_fn = clupoints_n_1_exp, proj_dist_fn = proj_beta)

plot_examples(list(e = e34, t = "e34: proj_dist_fn = 'norm' (default)"),
              list(e = e35, t = "e35: proj_dist_fn = 'unif'"),
              list(e = e36, t = "e36: custom proj_dist_fn (Beta dist.)"))
```

## Manipulating cluster sizes

```{r}
seed <- 87

# Custom clucenters_fn (all): yields fixed positions for the clusters
centers_fixed <- function(nclu, csep, coff) {
  matrix(c(-csep[1], -csep[2], csep[1], -csep[2], -csep[1], csep[2], csep[1], csep[2]),
         nrow = nclu, byrow = T)
}

# Custom clusizes_fn (e38): cluster sizes determined via the uniform distribution,
# no correction for total points
clusizes_unif <- function(nclu, npts, ae) sample(2 * npts / nclu, nclu, replace = T)

# Custom clusizes_fn (e39): clusters all have the same size, no correction for
# total points
clusizes_equal <- function(nclu, npts, ae) npts %/% nclu * rep.int(1, nclu)
```

```{r}
set.seed(seed)
e37 <- clugen(2, 4, 1500, c(1, 1), pi, c(20, 20), 0, 0, 5, 
             clucenters_fn = centers_fixed, point_dist_fn = "n")
set.seed(seed)
e38 <- clugen(2, 4, 1500, c(1, 1), pi, c(20, 20), 0, 0, 5,
             clucenters_fn = centers_fixed, clusizes_fn = clusizes_unif,
             point_dist_fn = "n")
set.seed(seed)
e39 <- clugen(2, 4, 1500, c(1, 1), pi, c(20, 20), 0, 0, 5,
             clucenters_fn = centers_fixed, clusizes_fn = clusizes_equal,
             point_dist_fn = "n")

plot_examples(list(e = e37, t = "e37: normal dist. (default)"),
              list(e = e38, t = "e38: unif. dist. (custom)"),
              list(e = e39, t = "e39: equal size (custom)"))
```
