---
title: "Merging and hierarchical cluster examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Merging and hierarchical cluster examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{ggplot2,patchwork}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  fig.asp = 0.5,
  dpi = 200
)
```

This section contains several examples on how to merge cluster data, either
generated with **clugenr** or from other sources. Although it is possible to
merge data in any dimension, these examples will focus on merging 2D data.
Therefore, we'll use the same `plot_examples_2d()` function used for the [2D examples](articles/examples3d2.html). Therefore, we require the following setup
code:

```{r setup, message=FALSE, warning=FALSE}
# Load the clugenr library
library(clugenr)

# Load functions for plotting examples
source("plot_examples_2d.R", local = knitr::knit_global())

# Keep examples reproducible in newer R versions
RNGversion("3.6.0")
```

## Merging two data sets generated with `clugen()`

```{r}
seed1 <- 4444
seed2 <- 5555
```

```{r}
e090 <- clugen(2, 5, 1000, c(1, 1), pi / 12, c(20, 20), 14, 1.2, 1.5, seed = seed1,
               proj_dist_fn = "unif", point_dist_fn = "n")
e091 <- clugen(2, 3, 1500, c(1, 0), 0.05, c(20, 20), 0, 0, 4, seed = seed2,
               point_dist_fn = "n", cluster_offset =c(20, 0))
e092 <- clumerge(e090, e091)
```

```{r}
plot_examples_2d(list(e = e090, t = "e090: data set 1"),
                 list(e = e091, t = "e091: data set 2"),
                 list(e = e092, t = "e092: merged data sets"))
```

In the previous example, clusters from individual data sets remain as separate
clusters in the merged data set. It's also possible to maintain the original
cluster labels by setting the `clusters_field` parameter to `NA`:

```{r}
e093 <- clumerge(e090, e091, clusters_field = NA)
```

```{r}
plot_examples_2d(list(e = e090, t = "e090: data set 1"),
                 list(e = e091, t = "e091: data set 2"),
                 list(e = e093, t = "e093: merged data sets"),
                 palette = "set2")
```

## Adding noise to a `clugen()`-generated data set

```{r}
set.seed(333)

e094 = list(points = matrix(120 * runif(2 * 500) - 60, ncol = 2),
            clusters = factor(rep(9L, 500)))     # Assign noise to cluster 9
e095 = clumerge(e094, e092, clusters_field = NA) # and keep it 9 when merging
```

```{r}
plot_examples_2d(list(e = e092, t = "e092: original merged data sets"),
                 list(e = e094, t = "e094: random uniform noise"),
                 list(e = e095, t = "e095: data sets with noise"),
                 palette = "set2", pmargin = -0.1)
```

## Merging with data not generated with `clugen()`

Data generated with [clugen] can be merged with other data sets, for example as
a way of augmenting them. In this example we perform [PCA][stats::prcomp] to
the [datasets::swiss] dataset to reduce its 6 variables to 2. As there are only
47 observations in this dataset, we'll use [clugen] to generate very similar
clusters and then use [clumerge] to create an augmented dataset:

```{r}
set.seed(888)
spc <- prcomp(swiss)
e096 <- list(points = spc$x[, 1:2], clusters = factor(rep(4L, dim(spc$x)[1])))
e097 <- clugen(2, 3, 50, c(-0.3, 1), 0., c(0, 0), 0, 0, 3,
               proj_dist_fn = "unif", clusizes_fn = c(20, 6, 25),
               clucenters_fn = matrix(c(-58, -5, 10, -40, 38, 0), byrow = TRUE, ncol = 2),
               angle_deltas_fn = c(0.02, -0.03, 0.1),
               llengths_fn = c(55, 100, 60))
e098 <- clumerge(e096, e097, clusters_field = NA)
```

```{r}
plot_examples_2d(list(e = e096, t = "e096: swiss PCA"),
                 list(e = e097, t = "e097: clugen-generated"),
                 list(e = e098, t = "e098: merged"),
                 palette = "seaborn")
```

We can also hierarchize clusters from different sources:

```{r}
e099 <- append(e096, list(hclusters = factor(rep(1L, dim(spc$x)[1]))))
e100 <- append(e097, list(hclusters = factor(rep(2L, length(e097$clusters)))))
e101 <- clumerge(e099, e100, fields = "points", clusters_field = "hclusters")
```

```{r}
plot_examples_2d(list(e = e099, t = "e099: swiss PCA 1 cluster"),
                 list(e = e100, t = "e100: clugen 1 cluster"),
                 list(e = e101, t = "e101: merged 2 clusters"),
                 clusters_field = "hclusters",
                 palette = "seaborn")
```

Note that this example is for illustration purposes only. Augmenting a dataset
this way is probably not a statistically sound approach for most problems.

