---
title: "Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette contains a number of examples on how to use `clugenr`. Each
example must be preceded with:

```{r setup, message=FALSE, warning=FALSE}
library(clugenr)
library(plotly)     # Required for plotting
RNGversion("3.6.0") # Keep examples reproducible in newer R versions 
```

## 2D examples

```{r}
marker2d <- list(size = 4, line = list(color = 'black', width = 0.2))
```

### Manipulating the direction of cluster-supporting lines

#### Using the `direction` parameter

```{r}
xaxis <- list(range = c(-10, 30))
yaxis <- list(scaleanchor = "x", scaleratio = 1, range = c(-22, 22))
seed <- 123
```

```{r}
set.seed(seed)
e01 <- clugen(2, 4, 200, c(1, 0), 0, c(10, 10), 10, 1.5, 0.5)
plot_ly(x = e01$points[, 1], y = e01$points[, 2], color = as.character(e01$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e01: direction = [1, 0]", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
set.seed(seed)
e02 <- clugen(2, 4, 200, c(1, 1), 0, c(10, 10), 10, 1.5, 0.5)
plot_ly(x = e02$points[, 1], y = e02$points[, 2], color = as.character(e02$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e02: direction = [1, 1]", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
set.seed(seed)
e03 <- clugen(2, 4, 200, c(0, 1), 0, c(10, 10), 10, 1.5, 0.5)
plot_ly(x = e03$points[, 1], y = e03$points[, 2], color = as.character(e03$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e03: direction = [0, 1]", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

#### Changing the `angle_disp` parameter and using a custom `angle_deltas_fn` function

```{r}
xaxis <- list(range = c(-38, 38))
yaxis <- list(scaleanchor = "x", scaleratio = 1, range = c(-20, 40))
seed <- 123
```

```{r}
set.seed(seed)
e04 <- clugen(2, 6, 500, c(1, 0), 0, c(10, 10), 10, 1.5, 0.5)
plot_ly(x = e04$points[, 1], y = e04$points[, 2], color = as.character(e04$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e04: angle_disp = 0", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```


```{r}
set.seed(seed)
e05 <- clugen(2, 6, 500, c(1, 0), pi / 8, c(10, 10), 10, 1.5, 0.5)
plot_ly(x = e05$points[, 1], y = e05$points[, 2], color = as.character(e05$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e05: angle_disp = π/8", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
# Custom angle_deltas function: arbitrarily rotate some clusters by 90 degrees
angdel_90_fn <- function(nclu, astd) sample(c(0, pi / 2), nclu, replace = T)

set.seed(seed)
e06 <- clugen(2, 6, 500, c(1, 0), 0, c(10, 10), 10, 1.5, 0.5, angle_deltas_fn = angdel_90_fn)
plot_ly(x = e06$points[, 1], y = e06$points[, 2], color = as.character(e06$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e06: custom angle_deltas function", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

### Manipulating the length of cluster-supporting lines

#### Using the `llength` parameter

```{r}
xaxis <- list() #range = c(-38, 38))
yaxis <- list(scaleanchor = "x", scaleratio = 1) #, range = c(-20, 40))
seed <- 321
```

```{r}
set.seed(seed)
e07 <- clugen(2, 5, 800, c(1, 0), pi / 10, c(10, 10), 0, 0, 0.5, point_dist_fn = "n")
plot_ly(x = e07$points[, 1], y = e07$points[, 2], color = as.character(e07$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e06: llength = 0", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
set.seed(seed)
e08 <- clugen(2, 5, 800, c(1, 0), pi / 10, c(10, 10), 10, 0, 0.5, point_dist_fn = "n")
plot_ly(x = e08$points[, 1], y = e08$points[, 2], color = as.character(e08$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e08: llength = 10", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
set.seed(seed)
e09 <- clugen(2, 5, 800, c(1, 0), pi / 10, c(10, 10), 30, 0, 0.5, point_dist_fn = "n")
plot_ly(x = e09$points[, 1], y = e09$points[, 2], color = as.character(e09$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e09: length = 30", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

#### Changing the `llength_disp` parameter and using a custom `llengths_fn` function

```{r}
xaxis <- list(range = c(-20, 29))
yaxis <- list(scaleanchor = "x", scaleratio = 1, range = c(-35, 20))
seed <- 765
```

```{r}
set.seed(seed)
e10 <- clugen(2, 5, 800, c(1, 0), pi / 10, c(10, 10), 15,  0.0, 0.5, point_dist_fn = "n")
plot_ly(x = e10$points[, 1], y = e10$points[, 2], color = as.character(e10$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e10: llength_disp = 0.0", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
set.seed(seed)
e11 <- clugen(2, 5, 800, c(1, 0), pi / 10, c(10, 10), 15, 10.0, 0.5, point_dist_fn = "n")
plot_ly(x = e11$points[, 1], y = e11$points[, 2], color = as.character(e11$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e11: llength_disp = 5.0", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
# Custom llengths function: line lengths grow for each new cluster
llen_grow_fn <- function(nclu, llen, llenstd) llen * 0:(nclu - 1) + rnorm(nclu, sd = llenstd)

set.seed(seed)
e12 <- clugen(2, 5, 800, c(1, 0), pi / 10, c(10, 10), 10,  0.1, 0.5, llengths_fn = llen_grow_fn, point_dist_fn = "n")
plot_ly(x = e12$points[, 1], y = e12$points[, 2], color = as.character(e12$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e12: custom llengths function", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

### Manipulating relative cluster positions

#### Using the `cluster_sep` parameter

```{r}
xaxis <- list(range = c(-130, 100))
yaxis <- list(scaleanchor = "x", scaleratio = 1, range = c(-100, 120))
seed <- 222
```

```{r}
set.seed(seed)
e13 <- clugen(2, 8, 1000, c(1, 1), pi / 4, c(10, 10), 15, 2, 4)
plot_ly(x = e13$points[, 1], y = e13$points[, 2], color = as.character(e13$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e13: cluster_sep = [10, 10]", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
set.seed(seed)
e14 <- clugen(2, 8, 1000, c(1, 1), pi / 4, c(30, 10), 15, 2, 4)
plot_ly(x = e14$points[, 1], y = e14$points[, 2], color = as.character(e14$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e14: cluster_sep = [30, 10]", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
set.seed(seed)
e15 <- clugen(2, 8, 1000, c(1, 1), pi / 4, c(10, 30), 15, 2, 4)
plot_ly(x = e15$points[, 1], y = e15$points[, 2], color = as.character(e15$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e15: cluster_sep = [10, 30]", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

#### Changing the `cluster_offset` parameter and using a custom `clucenters_fn` function

```{r}
xaxis <- list(range = c(-55, 55))
yaxis <- list(scaleanchor = "x", scaleratio = 1, range = c(-57, 45))
seed <- 222
```

```{r}
set.seed(seed)
e16 <- clugen(2, 8, 1000, c(1, 1), pi / 4, c(10, 10), 10, 2, 2.5)
plot_ly(x = e16$points[, 1], y = e16$points[, 2], color = as.character(e16$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e16: default", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
set.seed(seed)
e17 <- clugen(2, 8, 1000, c(1, 1), pi / 4, c(10, 10), 10, 2, 2.5,
              cluster_offset = c(20, -20))
plot_ly(x = e17$points[, 1], y = e17$points[, 2], color = as.character(e17$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e17: cluster_offset = [20, -20]", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
# Custom clucenters function: places clusters in a diagonal
centers_diag_fn <- function(nclu, csep, coff) {
  matrix(1, nrow = nclu, ncol = length(csep)) * (1:nclu * max(csep)) + rep(coff, each = nclu)
}

set.seed(seed)
e18 <- clugen(2, 8, 1000, c(1, 1), pi / 4, c(10, 10), 10, 2, 2.5, 
              cluster_offset = c(-50, -50), clucenters_fn = centers_diag_fn)
plot_ly(x = e18$points[, 1], y = e18$points[, 2], color = as.character(e18$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e18: custom clucenters function", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

### Lateral dispersion and placement of point projections on the line

```{r}
xaxis <- list(range = c(-30, 45))
yaxis <- list(scaleanchor = "x", scaleratio = 1, range = c(-60, 22))
seed <- 234
```

#### Normal projection placement (default): `proj_dist_fn = "norm"`

```{r}
set.seed(seed)
e19 <- clugen(2, 4, 1000, c(1, 0), pi / 2, c(20, 20), 13, 2, 0.0)
plot_ly(x = e19$points[, 1], y = e19$points[, 2], color = as.character(e19$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e19: lateral_disp = 0", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
set.seed(seed)
e20 <- clugen(2, 4, 1000, c(1, 0), pi / 2, c(20, 20), 13, 2, 1.0)
plot_ly(x = e20$points[, 1], y = e20$points[, 2], color = as.character(e20$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e20: lateral_disp = 1", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
set.seed(seed)
e21 <- clugen(2, 4, 1000, c(1, 0), pi / 2, c(20, 20), 13, 2, 3.0)
plot_ly(x = e21$points[, 1], y = e21$points[, 2], color = as.character(e21$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e21: lateral_disp = 3", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

#### Uniform projection placement: `proj_dist_fn = "unif"`

```{r}
set.seed(seed)
e22 <- clugen(2, 4, 1000, c(1, 0), pi / 2, c(20, 20), 13, 2, 0.0, proj_dist_fn = "unif")
plot_ly(x = e22$points[, 1], y = e22$points[, 2], color = as.character(e22$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e22: lateral_disp = 0", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
set.seed(seed)
e23 <- clugen(2, 4, 1000, c(1, 0), pi / 2, c(20, 20), 13, 2, 1.0, proj_dist_fn = "unif")
plot_ly(x = e23$points[, 1], y = e23$points[, 2], color = as.character(e23$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e23: lateral_disp = 1", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
set.seed(seed)
e24 <- clugen(2, 4, 1000, c(1, 0), pi / 2, c(20, 20), 13, 2, 3.0, proj_dist_fn = "unif")
plot_ly(x = e24$points[, 1], y = e24$points[, 2], color = as.character(e24$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e24: lateral_disp = 3", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

#### Custom projection placement using the Beta distribution

```{r}
# Custom proj_dist_fn: point projections placed using the Beta distribution
proj_beta <- function(len, n) len * rbeta(n, 0.1, 0.1) - len / 2
```

```{r}
set.seed(seed)
e25 <- clugen(2, 4, 1000, c(1, 0), pi / 2, c(20, 20), 13, 2, 0.0, proj_dist_fn = proj_beta)
plot_ly(x = e25$points[, 1], y = e25$points[, 2], color = as.character(e25$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e25: lateral_disp = 0", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
set.seed(seed)
e26 <- clugen(2, 4, 1000, c(1, 0), pi / 2, c(20, 20), 13, 2, 1.0, proj_dist_fn = proj_beta)
plot_ly(x = e26$points[, 1], y = e26$points[, 2], color = as.character(e26$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e26: lateral_disp = 1", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
set.seed(seed)
e27 <- clugen(2, 4, 1000, c(1, 0), pi / 2, c(20, 20), 13, 2, 3.0, proj_dist_fn = proj_beta)
plot_ly(x = e27$points[, 1], y = e27$points[, 2], color = as.character(e27$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e27: lateral_disp = 3", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

### Controlling final point positions from their projections on the cluster-supporting line

```{r}
xaxis <- list(range = c(-45, 31))
yaxis <- list(scaleanchor = "x", scaleratio = 1, range = c(-20, 40))
seed <- 12321
```

#### Points on hyperplane orthogonal to cluster-supporting line (default): `point_dist_fn = "n-1"`

```{r}
set.seed(seed)
e28 <- clugen(2, 5, 1500, c(1, 0), pi / 4, c(20, 20), 14, 2, 2.0)
plot_ly(x = e28$points[, 1], y = e28$points[, 2], color = as.character(e28$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e28: proj_dist_fn=\"norm\" (default)", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
set.seed(seed)
e29 <- clugen(2, 5, 1500, c(1, 0), pi / 4, c(20, 20), 14, 2, 2.0, proj_dist_fn = "unif")
plot_ly(x = e29$points[, 1], y = e29$points[, 2], color = as.character(e29$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e29: proj_dist_fn=\"unif\"", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
# Custom proj_dist_fn: point projections placed using the Beta distribution
proj_beta <- function(len, n) len * rbeta(n, 0.1, 0.1) - len / 2

set.seed(seed)
e30 <- clugen(2, 5, 1500, c(1, 0), pi / 4, c(20, 20), 14, 2, 2.0, proj_dist_fn = proj_beta)
plot_ly(x = e30$points[, 1], y = e30$points[, 2], color = as.character(e30$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e30: custom proj_dist_fn (Beta)", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

#### Points around projection on cluster-supporting line: `point_dist_fn = "n"`

```{r}
set.seed(seed)
e31 <- clugen(2, 5, 1500, c(1, 0), pi / 4, c(20, 20), 14, 2, 2.0,
              point_dist_fn = "n")
plot_ly(x = e31$points[, 1], y = e31$points[, 2], color = as.character(e31$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e31: proj_dist_fn=\"norm\" (default)", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
set.seed(seed)
e32 <- clugen(2, 5, 1500, c(1, 0), pi / 4, c(20, 20), 14, 2, 2.0,
              point_dist_fn = "n", proj_dist_fn = "unif")
plot_ly(x = e32$points[, 1], y = e32$points[, 2], color = as.character(e32$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e32: proj_dist_fn=\"unif\"", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
# Custom proj_dist_fn: point projections placed using the Beta distribution
proj_beta <- function(len, n) len * rbeta(n, 0.1, 0.1) - len / 2

set.seed(seed)
e33 <- clugen(2, 5, 1500, c(1, 0), pi / 4, c(20, 20), 14, 2, 2.0,
              point_dist_fn = "n", proj_dist_fn = proj_beta)
plot_ly(x = e33$points[, 1], y = e33$points[, 2], color = as.character(e33$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e33: custom proj_dist_fn (Beta)", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

#### Custom point placement using the exponential distribution

```{r}
# Custom point_dist_fn: final points placed using the Exponential distribution
clupoints_n_1_exp <- function(projs, lat_std, len, clu_dir, clu_ctr) {
    dist_exp <- function(npts, lstd) lstd * rexp(npts, rate = 2 / lstd)
    clupoints_n_1_template(projs, lat_std, clu_dir, dist_exp)
}
```

```{r}
set.seed(seed)
e34 <- clugen(2, 5, 1500, c(1, 0), pi / 4, c(20, 20), 14, 2, 2.0,
              point_dist_fn = clupoints_n_1_exp)
plot_ly(x = e34$points[, 1], y = e34$points[, 2], color = as.character(e34$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e34: proj_dist_fn=\"norm\" (default)", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
set.seed(seed)
e35 <- clugen(2, 5, 1500, c(1, 0), pi / 4, c(20, 20), 14, 2, 2.0,
              point_dist_fn = clupoints_n_1_exp, proj_dist_fn = "unif")
plot_ly(x = e35$points[, 1], y = e35$points[, 2], color = as.character(e35$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e35: proj_dist_fn=\"unif\"", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
# Custom proj_dist_fn: point projections placed using the Beta distribution
proj_beta <- function(len, n) len * rbeta(n, 0.1, 0.1) - len / 2

set.seed(seed)
e36 <- clugen(2, 5, 1500, c(1, 0), pi / 4, c(20, 20), 14, 2, 2.0,
              point_dist_fn = clupoints_n_1_exp, proj_dist_fn = proj_beta)
plot_ly(x = e36$points[, 1], y = e36$points[, 2], color = as.character(e36$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e36: custom proj_dist_fn (Beta)", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

### Manipulating cluster sizes

```{r}
xaxis <- list(range = c(-40, 40))
yaxis <- list(scaleanchor = "x", scaleratio = 1, range = c(-40, 40))
seed <- 87

# Custom clucenters_fn (all): yields fixed positions for the clusters
centers_fixed <- function(nclu, csep, coff) {
  matrix(c(-csep[1], -csep[2], csep[1], -csep[2], -csep[1], csep[2], csep[1], csep[2]),
         nrow = nclu, byrow = T)
}
```

```{r}
set.seed(seed)
e37 <- clugen(2, 4, 1500, c(1, 1), pi, c(20, 20), 0, 0, 5, 
             clucenters_fn = centers_fixed, point_dist_fn = "n")
plot_ly(x = e37$points[, 1], y = e37$points[, 2], color = as.character(e37$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e37: normal dist. (default)", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
# Custom clusizes_fn (e38): cluster sizes determined via the uniform distribution, no correction for total points
clusizes_unif <- function(nclu, npts, ae) sample(2 * npts / nclu, nclu, replace = T)

set.seed(seed)
e38 <- clugen(2, 4, 1500, c(1, 1), pi, c(20, 20), 0, 0, 5,
             clucenters_fn = centers_fixed, clusizes_fn = clusizes_unif, point_dist_fn = "n")
plot_ly(x = e38$points[, 1], y = e38$points[, 2], color = as.character(e38$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e38: unif. dist. (custom)", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

```{r}
# Custom clusizes_fn (e39): clusters all have the same size, no correction for total points
clusizes_equal <- function(nclu, npts, ae) npts %/% nclu * rep.int(1, nclu)

set.seed(seed)
e39 <- clugen(2, 4, 1500, c(1, 1), pi, c(20, 20), 0, 0, 5,
             clucenters_fn = centers_fixed, clusizes_fn = clusizes_equal, point_dist_fn = "n")
plot_ly(x = e39$points[, 1], y = e39$points[, 2], color = as.character(e39$point_clusters),
        type = "scatter", mode = "markers", marker = marker2d) %>%
  layout(title = "e39: equal size (custom)", xaxis = xaxis, yaxis = yaxis, showlegend = F)
```

## 3D examples

```{r}
marker3d <- list(size = 3, line = list(color = 'black', width = 0.3))
```

### Manipulating the direction of cluster-supporting lines

#### Using the `direction` parameter

```{r}
scene <- list(xaxis = list(range = list(-12, 27), dtick = 10),
              yaxis = list(range = list(-17, 22), dtick = 10),
              zaxis = list(range = list(-22, 22), dtick = 10),
              aspectmode = "cube")
seed <- 123
```

```{r}
set.seed(seed)
e40 <- clugen(3, 4, 500, c(1, 0, 0), 0, c(10, 10, 10), 15, 1.5, 0.5)
plot_ly(x = e40$points[, 1], y = e40$points[, 2], z = e40$points[, 3],
        color = as.character(e40$point_clusters),
        type = "scatter3d", mode = "markers", marker = marker3d) %>%
  layout(title = "e40: direction = [1, 0, 0]", showlegend= F, scene = scene)
```

```{r}
set.seed(seed)
e41 <- clugen(3, 4, 500, c(1, 1, 1), 0, c(10, 10, 10), 15, 1.5, 0.5)
plot_ly(x = e41$points[, 1], y = e41$points[, 2], z = e41$points[, 3],
        color = as.character(e41$point_clusters),
        type = "scatter3d", mode = "markers", marker = marker3d) %>%
  layout(title = "e41: direction = [1, 1, 1]", showlegend= F, scene = scene)
```

```{r}
set.seed(seed)
e42 <- clugen(3, 4, 500, c(0, 0, 1), 0, c(10, 10, 10), 15, 1.5, 0.5)
plot_ly(x = e42$points[, 1], y = e42$points[, 2], z = e42$points[, 3],
        color = as.character(e42$point_clusters),
        type = "scatter3d", mode = "markers", marker = marker3d) %>%
  layout(title = "e42: direction = [0, 0, 1]", showlegend= F, scene = scene)
```

#### Changing the `angle_disp` parameter and using a custom `angle_deltas_fn` function

```{r}
scene <- list(xaxis = list(range = list(-35, 30), dtick = 10),
              yaxis = list(range = list(-15, 40), dtick = 10),
              zaxis = list(range = list(-25, 17), dtick = 10),
              aspectmode = "cube")
seed <- 123
```

```{r}
set.seed(seed)
e43 <- clugen(3, 6, 1000, c(1, 0, 0), 0, c(10, 10, 10), 15, 1.5, 0.5)
plot_ly(x = e43$points[, 1], y = e43$points[, 2], z = e43$points[, 3],
        color = as.character(e43$point_clusters),
        type = "scatter3d", mode = "markers", marker = marker3d) %>%
  layout(title = "e43: angle_disp = 0", showlegend = F, scene = scene)
```

```{r}
set.seed(seed)
e44 <- clugen(3, 6, 1000, c(1, 0, 0), pi / 8, c(10, 10, 10), 15, 1.5, 0.5)
plot_ly(x = e44$points[, 1], y = e44$points[, 2], z = e44$points[, 3],
        color = as.character(e44$point_clusters),
        type = "scatter3d", mode = "markers", marker = marker3d) %>%
  layout(title = "e44: angle_disp = π/8", showlegend = F, scene = scene)
```

```{r}
# Custom angle_deltas function: arbitrarily rotate some clusters by 90 degrees
angdel_90 <- function(nclu, astd) sample(c(0, pi / 2), nclu, replace = T)

set.seed(seed)
e45 <- clugen(3, 6, 1000, c(1, 0, 0), 0, c(10, 10, 10), 15, 1.5, 0.5, angle_deltas_fn = angdel_90)
plot_ly(x = e45$points[, 1], y = e45$points[, 2], z = e45$points[, 3],
        color = as.character(e45$point_clusters),
        type = "scatter3d", mode = "markers", marker = marker3d) %>%
  layout(title = "e45: custom angle_deltas function", showlegend = F, scene = scene)
```



